Robust receding horizon control for linear systems with state and input dependent disturbances

### 2015-06-29 example.tex
Added label to figure.
Post-review edits

%!TEX root = draft_v2.tex
\section{Complexity}\label{sec:complexity}
We next discuss bounds on the computational complexity of the algorithm outlined in Section~\ref{sec:line:search}.
For a given active set at most $2N$ mpQPs must be solved, and this is done by decomposing the
problem as stated in section~\ref{sec:gen:mpQP}.  For each mpQP the main computational effort is to
determine $\Sigma$ in~\eqref{seq:sigma}. In the case of~\eqref{eq:sigma:normal} this requires a QR
decomposition with $2(mn^2-n^3/3)$ floating point operations (flops, see e.g.~\cite{Golub:1996}), 
where $m$ is the number of active constraints and $n$ the number of
optimisation variables, plus inversion of a triangular matrix with $(m^3-m)/6$ flops and several matrix
multiplications with $mpn$ flops (for the multiplication of a $m\times n$ matrix with a $n\times p$
matrix). The computationally more expensive case of~\eqref{eq:sigma:degen} requires a QR decomposition,
inversion of a definite matrix $W$ using Cholesky decomposition with $n^3/3$ flops, inversion of a triangular
matrix, several matrix multiplications and additions with $mn$ flops, and inversion of a definite matrix
$\Delta$. In the worst case this adds up to at most $11n^3+n^2-n/3$ flops to obtain $\Sigma$. The solution to
the problem is then the result of a matrix multiplication, and, in order to be able to solve the subsequent
mpQP, the matrix $W_\theta$ and the vector $c_\theta$ also need to be computed. Hence the overall flop count
in the worst case adds up to $11n^3+n^2+2/3n+2n^2 n_\theta$, where $n_\theta$ denotes the dimension of the
parameter~$\theta$. 
%At each stage two mpQPs have to be solved recursively, so that 
In the worst case the
recursion for the entire horizon therefore requires $N(2 n_x+5 n_x^2+101 n_x^3+8 n_x^2 (n_u+n_x))$ flops.

@@

\section{Example}\label{sec:example}
In this section we present a numerical simulation for a simplified levitating ball model, as derived in~\cite{Schaich:2015}. The dynamics of the nonlinear system are given $m\ddot y = m g - c\frac{i^2}{y^2}$,
where $m$ denotes the mass of the ball, $g$ denotes the gravitation constant, $c$ denotes a constant depending
on the coil, $y$ the vertical distance to the coil, the simplification is due to neglecting inductive dynamics 
and assuming we that can influence the current~$i$ directly. Figure~\ref{fig:levitating:ball} illustrates the setup.
Choosing the state $x=(y,\dot y)$ and the input $u=i$ we obtain a state space representation which depends nonlinearly
on both the state and the input. For any $x_1>0$, the system is in equilibrium if $x_2=0$ and $u=\sqrt{\frac{gm}{c}}x_1$.

Linearising the nonlinear differential 
equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
%
\[
	 \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
	0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
	\end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
	0 \\ - \frac{2c\hat u}{m\hat x_1^2}
	\end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
\]
%
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using
the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving
\[
\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}
\]
%
A representation of the form~\eqref{eq:the:problem}, in which the disturbance $w$ accounts for linearisation
errors, can be obtained by sampling input and state
constraints and using the mean-value theorem \citep[see][]{Schaich:2015}. This leads
to a disturbance representation of the form $\mathcal W(x,u)=\{w\in\mathbb R^n: 
\min_{k\in\mathcal S_w}\{ H_{k,i}^x x + H_{k,i}^u u\} \leq w_i \leq \max_{k\in\mathcal S_w} 
\{ H_{k,i}^x x + H_{k,i}^u u \} \}$,
where ${k\in\mathcal S_w}$ denotes the sample index and $i$ the row number. This point-wise polytopic representation
together with a polyhedral representation of state constraints allows us to compute a polyhedral
MRPI set~$\mathcal X^\infty$. As a terminal feedback controller $K$ we use the solution to the robust 
Lyapunov condition $V(x)-V((A+BK)x+Dw)\leq \gamma^2 w^Tw$ 
with $V(x)=x^T P_0 x\geq0$, i.e. $x^TP_0x - ((A+BK)x+Dw)^TP_0((A+BK)x+Dw)\geq x^T(Q+K^TRK)x -\gamma^2 w^T w$, 
\cite[see e.g.][]{Boyd:94}. This semi-definite program can also be used to find the minimum value of $\gamma^2$. Starting from 
a polyhedral MRPI set~$\mathcal X^\infty$ we apply the algorithm presented in section~\ref{sec:preliminaries}
to obtain the necessary stage constraints~$\mathcal Z_k$. At this stage we can introduce numerical values
for individual parameters: $T_s=30\,\text{ms}$, $C=1\,\frac{\text{kg m}^3}{\text{A}^2\text{s}^2}$,
$m=100\,\text{g}$, $\hat x_1 = 50\,\text{mm}$ and $\mathcal X=\{x:
\abs{x_1- \hat x_1}\leq 1\,\text{mm}\wedge \abs{x_2}\leq 105\,\frac{\text{mm}}{\text{s}}\}$, $\mathcal
U=\{u:\abs{ u-\hat u}\leq 10 \,\text{mA}\}$.
To obtain $\mathcal W(x,u)$ we sample $\mathcal X$ and $\mathcal U$ with a total of 25 samples, i.e. the cardinality
of $\mathcal S_w$ is 10. In the simulation we use horizon length $N=5$. We perform a cold started line search from $\chi_0=(0,0)$
to the outer bound of the constraint set at $\chi_e=(1,-30)$, the resulting trajectories at the boundaries of
individual active sets are plotted in figure~\ref{fig:plot}.

The simulation was implemented in MATLAB only using non-optimised built-in functions for the main computational
purposes. In the simulation illustrated in figure~\ref{fig:plot} there were a total of 10 iterations required,
out of which 7 correspond to changes of the dominating right hand side of $\mathcal W(x,u)$, for which no
recomputation of the solution is necessary. For $\chi_e$ on the boundary of the feasible set
the solution time is less than $0.5\, \text{s}$ in total (using a MacBook Pro with a $2.3\,\text{GHz}$ quad-core processor).

>>>>

%!TEX root = post_review.tex
\section{Complexity}\label{sec:complexity}
We next discuss bounds on the computational complexity of the algorithm outlined in Section~\ref{sec:line:search}.
For a given active set at most $2N$ mpQPs must be solved, and this is done by decomposing the
problem as stated in section~\ref{sec:gen:mpQP}.  For each mpQP the main computational effort is to
determine $\Sigma$ in~\eqref{seq:sigma}. In the case of~\eqref{eq:sigma:normal} this requires a QR
decomposition with $2(mn^2-n^3/3)$ floating point operations (flops, see e.g.~\cite{Golub:1996}), 
where $m$ is the number of active constraints and $n$ the number of
optimisation variables, plus inversion of a triangular matrix with $(m^3-m)/6$ flops and several matrix
multiplications with $mpn$ flops (for the multiplication of a $m\times n$ matrix with a $n\times p$
matrix). The computationally more expensive case of~\eqref{eq:sigma:degen} requires a QR decomposition,
inversion of a definite matrix $W$ using Cholesky decomposition with $n^3/3$ flops, inversion of a triangular
matrix, several matrix multiplications and additions with $mn$ flops, and inversion of a definite matrix
$\Delta$. In the worst case this adds up to at most $11n^3+n^2-n/3$ flops to obtain $\Sigma$. The solution to
the problem is then the result of a matrix multiplication, and, in order to be able to solve the subsequent
mpQP, the matrix $W_\theta$ and the vector $c_\theta$ also need to be computed. Hence the overall flop count
in the worst case adds up to $11n^3+n^2+2/3n+2n^2 n_\theta$, where $n_\theta$ denotes the dimension of the
parameter~$\theta$. 
%At each stage two mpQPs have to be solved recursively, so that 
In the worst case the
recursion for the entire horizon therefore requires $N(2 n_x+5 n_x^2+101 n_x^3+8 n_x^2 (n_u+n_x))$ flops.

@@

\section{Example}\label{sec:example}
In this section we present a numerical simulation for a simplified levitating ball model, as derived in~\cite{Schaich:2015}. \edit{The dynamics of the nonlinear system are given $m\ddot y = m g - c i^2/y^2$,
where $m$ denotes the mass of the ball, $g$ is the acceleration due to gravity, $c$ is a constant depending
on the coil, $y$ the vertical distance to the coil (Fig.~\ref{fig:levitating:ball}). We assume that the current~$i$ is directly specified by the controller, thus neglecting the inductive dynamics of the coil. %Figure~\ref{fig:levitating:ball} illustrates the setup.
Choosing the state $x=(y,\dot y)$ and the input $u=i$ we obtain a state space representation which depends nonlinearly
on the state and control input.} For any $x_1>0$, the system is in equilibrium if $x_2=0$ and $u=\sqrt{\frac{gm}{c}}x_1$.

Linearising the nonlinear differential 
equation $\dot x = f(x,u)$ around an equilibrium point $(\hat x, \hat
u)$ gives the approximate linear model 
%
\[
	 \Delta\dot{x} = \underbrace{\left(\begin{array}{cc}
	0 & 1 \\ \frac{2c\hat u^2}{m\hat x_1^3} & 0
	\end{array}\right)}_{\frac{\partial f}{\partial x}(\hat x,\hat
      u)}\Delta x 
+ \underbrace{\left(\begin{array}{c}
	0 \\ - \frac{2c\hat u}{m\hat x_1^2}
	\end{array}\right)}_{\frac{\partial f}{\partial u}(\hat x,\hat
      u)}\Delta u
\]
%
where $\Delta u = u -\hat{u}$ and $\Delta x \approx x-\hat{x}$.
We derive the discrete time dynamics with sampling rate $T_s$ using
the Euler formula $x^+=x+T_s f(x,u) =:\tilde f(x,u)$ giving
\[
\Delta x^+ = A \Delta x + B \Delta u , \quad
\Biggl\{\begin{aligned} A &= I+T_s\frac{\partial f}{\partial  x}(\hat
  x,\hat u) \\
B &= T_s \frac{\partial f}{\partial u}(\hat x,\hat u)
\end{aligned}
\]
%
A representation of the form~\eqref{eq:the:problem}, in which the disturbance $w$ accounts for linearisation
errors, can be obtained by sampling input and state
constraints and using the mean-value theorem \citep[see][]{Schaich:2015}. This leads
to a disturbance representation of the form $\mathcal W(x,u)=\{w\in\mathbb R^n: 
\min_{k\in\mathcal S_w}\{ H_{k,i}^x x + H_{k,i}^u u\} \leq w_i \leq \max_{k\in\mathcal S_w} 
\{ H_{k,i}^x x + H_{k,i}^u u \} \}$,
where ${k\in\mathcal S_w}$ denotes the sample index and $i$ the row number. This point-wise polytopic representation
together with a polyhedral representation of state constraints allows us to compute a polyhedral
MRPI set~$\mathcal X^\infty$. As a terminal feedback controller $K$ we use the solution to the robust 
Lyapunov condition $V(x)-V((A+BK)x+Dw)\leq \gamma^2 w^Tw$ 
with $V(x)=x^T P_0 x\geq0$, i.e. $x^TP_0x - ((A+BK)x+Dw)^TP_0((A+BK)x+Dw)\geq x^T(Q+K^TRK)x -\gamma^2 w^T w$, 
\cite[see e.g.][]{Boyd:94}. This semi-definite program can also be used to find the minimum value of $\gamma^2$. Starting from 
a polyhedral MRPI set~$\mathcal X^\infty$ we apply the algorithm presented in section~\ref{sec:preliminaries}
to obtain the necessary stage constraints~$\mathcal Z_k$. At this stage we can introduce numerical values
for individual parameters: $T_s=30\,\text{ms}$, $C=1\,\frac{\text{kg m}^3}{\text{A}^2\text{s}^2}$,
$m=100\,\text{g}$, $\hat x_1 = 50\,\text{mm}$ and $\mathcal X=\{x:
\abs{x_1- \hat x_1}\leq 1\,\text{mm}\wedge \abs{x_2}\leq 105\,\frac{\text{mm}}{\text{s}}\}$, $\mathcal
U=\{u:\abs{ u-\hat u}\leq 10 \,\text{mA}\}$.
To obtain $\mathcal W(x,u)$ we sample $\mathcal X$ and $\mathcal U$ with a total of 25 samples, i.e. the cardinality
of $\mathcal S_w$ is 10. In the simulation we use horizon length $N=5$. We perform a cold started line search from $\chi_0=(0,0)$
to the outer bound of the constraint set at $\chi_e=(1,-30)$, the resulting trajectories at the boundaries of
individual active sets are plotted in Figure~\ref{fig:plot}.

The simulation was implemented in MATLAB using only non-optimised built-in functions for the main computational
purposes. In the simulation illustrated in Fig.~\ref{fig:plot} a total of $10$ iterations were required,
of which $7$ correspond to changes of the dominating right hand side of $\mathcal W(x,u)$, for which no
recomputation of the solution is necessary. For $\chi_e$ on the boundary of the feasible set
the solution time is less than $0.5\, \text{s}$ in total (using a MacBook Pro with a $2.3\,\text{GHz}$ quad-core processor).
